\bf Comparing the Power of Plot Designs to Reveal Correlation

### 2015-12-10 paper/paper.Rnw
its working, with references

<<setup, echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
library(lme4)
library(ggplot2)
library(mnormt)
library(tidyr)
# Convert reshape2 to tidyr, melt to gather - done
#library(reshape2)
library(nullabor)
library(dplyr)
library(mFilter)
library(grid)
library(gridExtra)
#library(splitstackshape) - changed cSplit to unnest
#library(readr)
#library(lubridate)
# Convert these to read_csv at some point
correl <- read.csv("correlations.csv", stringsAsFactors = FALSE)
full <- read.csv("full.csv", row.names=1, stringsAsFactors = FALSE) 
Turk <- read.csv("turk.csv", stringsAsFactors = FALSE)

@@

\begin{figure}[htbp]
 \centering
<<results, dependson='setup', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE>>=
Turk <- mutate(Turk, plot = ifelse(test_param =="line", "Overlaid Lines", "Scatterplot"), 
               sign=ifelse(correlation > 0, 1, 0), smoothed=ifelse(smoothed=="yes", 1, 0))
#ggplot(data=subset(Turk, smoothed==0), aes(x=correlation, y=percent, colour=plot)) + geom_smooth() + geom_point() + facet_grid(n~.) + scale_colour_discrete(name = "Plot Design") +  labs(title="Unsmoothed Plot Accuracy", y="Percent Correct", x="True Plot Correlation")
ggplot(data=Turk, aes(x=correlation, y=percent, colour=plot)) + geom_smooth(se=F) + geom_point() + facet_grid(n~smoothed) + 
  scale_colour_discrete(name = "Plot Design") + labs(title="Observed Data", y="Percent Correct", x="Correlation") +
  theme(legend.position="bottom")
@
  \label{results}
  \caption{Response accuracy for unsmoothed data, broken by unsmoothed (left), smoothed (right) and sample size. The dots are data points, the lines represent a naive model average fit. Scatter plots generally showed superior performance, with the overlaid line graphs being particularly weak at revealing negative correlation}

@@

\begin{figure}[htbp]
  \centering
<<powerdiff, dependson='results', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE>>=
Turk$correct <- Turk$percent*Turk$attempt
Turk$r <- round(Turk$correlation, 1)
Turk.line <- subset(Turk, test_param=="line")
Turk.scatter <- subset(Turk, test_param=="scatter")
Turk.line <- Turk.line[order(Turk.line$r, Turk.line$n, Turk.line$smoothed),c(10, 13, 16, 19, 20)]
Turk.line <- summarise(group_by(Turk.line, smoothed, n, r), sum(attempt), sum(correct))
Turk.scatter <- Turk.scatter[order(Turk.scatter$r, Turk.scatter$n, Turk.scatter$smoothed),c(10, 13, 16, 19, 20)]
Turk.scatter <- summarise(group_by(Turk.scatter, smoothed, n, r), sum(attempt), sum(correct))
colnames(Turk.scatter) <- c("smoothed", "n", "r", "s.attempt", "s.correct")
colnames(Turk.line) <- c("smoothed", "n", "r", "l.attempt", "l.correct")
prop <- cbind(Turk.line, Turk.scatter[4:5])
prop <- mutate(prop, l.adj = (l.correct+1)/(l.attempt+1), s.adj = (s.correct+1)/(s.attempt+1),
               df = s.attempt + l.attempt - 1, difference=s.adj-l.adj, t = qt(0.975, df),
               upper = difference + t*sqrt(s.adj*(1-s.adj)/s.attempt+l.adj*(1-l.adj)/l.attempt),
               lower = difference - t*sqrt(s.adj*(1-s.adj)/s.attempt+l.adj*(1-l.adj)/l.attempt))
ggplot(data=prop) + geom_line(aes(x=r, y=difference), colour="#99CC99", size=2) + 
                    geom_ribbon(aes(x=r, ymin=lower, ymax=upper), colour="#99CC99", fill="#99CC99", alpha=0.2) + 
                    labs(title="Difference in Power", x="Correlation") + facet_grid(n~smoothed)
@
 \label{powerdiff}
 \caption{Difference in estimated power to reveal the true plot by smoothness and sample size with a 95\% confidence interval. Values above zero indicate the scatter plot is more powerful than the overlaid line graph.}
\end{figure

@@

>>>>

<<setup, echo=FALSE, results='hide', message=FALSE, error=FALSE, warning=FALSE>>=
library(lme4)
library(ggplot2)
library(mnormt)
library(tidyr)
# Convert reshape2 to tidyr, melt to gather - done
#library(reshape2)
library(nullabor)
library(dplyr)
library(mFilter)
library(grid)
library(gridExtra)
library(stringr)
#library(splitstackshape) - changed cSplit to unnest
#library(readr)
#library(lubridate)
# Convert these to read_csv at some point
correl <- read.csv("correlations.csv", stringsAsFactors = FALSE)
full <- read.csv("full.csv", row.names=1, stringsAsFactors = FALSE) 
Turk <- read.csv("turk.csv", stringsAsFactors = FALSE)

@@

\begin{figure}[htbp]
 \centering
<<results, dependson='setup', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE>>=
Turk <- mutate(Turk, plot = ifelse(test_param =="line", "Overlaid Lines", "Scatterplot"), 
               sign=ifelse(correlation > 0, 1, 0), smoothed=ifelse(smoothed=="yes", 1, 0))
#ggplot(data=subset(Turk, smoothed==0), aes(x=correlation, y=percent, colour=plot)) + geom_smooth() + geom_point() + facet_grid(n~.) + scale_colour_discrete(name = "Plot Design") +  labs(title="Unsmoothed Plot Accuracy", y="Percent Correct", x="True Plot Correlation")
ggplot(data=Turk, aes(x=correlation, y=percent, colour=plot)) + geom_smooth(se=F) + geom_point() + facet_grid(n~smoothed) + 
  scale_colour_discrete(name = "Plot Design") + labs(title="Observed Data", y="Percent Correct", x="Correlation") +
  theme(legend.position="bottom")
@
  \label{results}
  \caption{Response accuracy for unsmoothed data, broken by unsmoothed (left), smoothed (right) and sample size. The dots are data points, the lines represent a naive model average fit. Scatter plots generally showed superior performance, with the overlaid line graphs being particularly weak at revealing negative correlation}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
<<powerdiff, dependson='results', fig.width=6, fig.height=8, out.width='0.7\\textwidth', echo=FALSE>>=
Turk$correct <- Turk$percent*Turk$attempt
Turk$r <- round(Turk$correlation, 1)
Turk.line <- subset(Turk, test_param=="line")
Turk.scatter <- subset(Turk, test_param=="scatter")
Turk.line <- Turk.line[order(Turk.line$r, Turk.line$n, Turk.line$smoothed),c(10, 13, 16, 19, 20)]
Turk.line <- summarise(group_by(Turk.line, smoothed, n, r), sum(attempt), sum(correct))
Turk.scatter <- Turk.scatter[order(Turk.scatter$r, Turk.scatter$n, Turk.scatter$smoothed),c(10, 13, 16, 19, 20)]
Turk.scatter <- summarise(group_by(Turk.scatter, smoothed, n, r), sum(attempt), sum(correct))
colnames(Turk.scatter) <- c("smoothed", "n", "r", "s.attempt", "s.correct")
colnames(Turk.line) <- c("smoothed", "n", "r", "l.attempt", "l.correct")
prop <- cbind(Turk.line, Turk.scatter[4:5])
prop <- mutate(prop, l.adj = (l.correct+1)/(l.attempt+1), s.adj = (s.correct+1)/(s.attempt+1),
               df = s.attempt + l.attempt - 1, difference=s.adj-l.adj, t = qt(0.975, df),
               upper = difference + t*sqrt(s.adj*(1-s.adj)/s.attempt+l.adj*(1-l.adj)/l.attempt),
               lower = difference - t*sqrt(s.adj*(1-s.adj)/s.attempt+l.adj*(1-l.adj)/l.attempt))
ggplot(data=prop) + geom_line(aes(x=r, y=difference), colour="#99CC99", size=2) + 
                    geom_ribbon(aes(x=r, ymin=lower, ymax=upper), colour="#99CC99", fill="#99CC99", alpha=0.2) + 
                    labs(title="Difference in Power", x="Correlation") + facet_grid(n~smoothed)
@
 \label{powerdiff}
 \caption{Difference in estimated power to reveal the true plot by smoothness and sample size with a 95\% confidence interval. Values above zero indicate the scatter plot is more powerful than the overlaid line graph.}
\end{figure}

@@

Here's how to do references \citet{hadley:2009} and \citep{hadley:2009}.

\bibliographystyle{asa}
\bibliography{references}
